How One Microtask Affects Another

### 2015-09-25 docs/publications/2015.08_CHI/man.tex
Revised manuscript

\numberofauthors{3}
\author{%
  \alignauthor{
	Edward Newell \\
    \affaddr{School of Computer Science, McGill University}\\
    \affaddr{3480 University Street, Room 318, Montreal, Qu\'ebec,
	  Canada, H3A 2A7
	}\\
	\email{edward.newell@mail.mcgill.ca}}\\
  \alignauthor{
	Derek Ruths \\
    \affaddr{School of Computer Science, McGill University}\\
    \affaddr{3480 University Street, Room 318, Montreal, Qu\'ebec,
	  Canada, H3A 2A7
	}\\
	\email{derek.ruths@mcgill.ca}
  }\\
}

@@

\begin{abstract}
Microtask platforms are becoming commonplace tools for performing human
research, producing gold-standard data, and annotating large datasets.
These platforms connect \textit{requesters}
(researchers or companies) with large populations (crowds) of workers, who 
perform small tasks, typically taking less than five minutes each.
A topic of ongoing research concerns the design of tasks that elicit high
quality annotations.
Here we identify a feature of nearly all crowdsourcing 
workflows that profoundly impacts workers' responses.
Microtask assignments typically consist of a sequence 
of tasks sharing a common format (e.g., circle galaxies in an image). 
Using image-labeling, a canonical microtask format, we 
discover that earlier tasks have a priming effect on the worker, shifting 
the distribution of future responses by 30-50\% 
(total variational distance). 
Specifically, prior tasks influence the content that workers focus on, 
as well as the richness and specialization of responses. 
We call this phenomenon \textit{intertask effects}.
We compare intertask effects to the overt framing, 
effected by stating the requester's research interest, 
and find that intertask effects are on par or stronger.
While intertask effects can be a source of systematic bias, 
our results suggest that, with appropriate task design, 
they might be leveraged to hone worker focus and acuity, 
helping to elicit reproducible, expert-level judgments.
Intertask effects are a crucial aspect of human computation that should be
considered in the design of any crowdsourced study.
\end{abstract}

@@

\section{Introduction}
There are many tasks that are trivial for people, but difficult to solve
programmatically.  
Typical tasks include tagging and categorizing images,
%\cite{6116320,Zhai2012357}, 
coding and transcribing media,
%\cite{chandler2013breaking,Berinsky2012351,Finnerty2013},%paolacci2010running},
%judging the relevancy or quality of content,
and performing surveys for 
academic or market research purposes (see Table~\ref{table:task_composition} for a listing of task types seen on the Amazon's Mechanical Turk 
microtask platform).
%\cite{le2010ensuring,grady2010crowdsourcing,alonso2009can,kazai2013analysis}.
Many tasks are ill-defined, in the sense that they do not have
a clear ``correct'' response, and require high-level, qualitative judgment.
Microtask platforms are marketplaces that help fill the gap in current 
computational capabilities, by matching requesters, who need to have such 
tasks completed, with human workers.  Amazon's Mechanical Turk\footnote{\url{mturk.com}}, and 
CrowdFlower\footnote{\url{crowdflower.com}} are two popular microtask 
platforms.

@@

To investigate the phenomenon in a
highly generalizable setting, we sought a task that would involve 
qualitative judgment and enable relatively unconstrained responses. 
We also sought a task that would be a typical exemplar of the kinds of
tasks actually seen on microtask platforms (see 
Table~\ref{table:task_composition} for examples).  For this purpose, we 
adopted
image-labeling as a canonical microtask, in which workers provide 
descriptive labels to images using free-text inputs.  Image labeling is
a qualitative task without clearly ``correct'' or ``incorrect'' responses,
and is among the most common kinds of tasks on the major crowdsourcing
platform Amazon's Mechanical Turk.  If task-induced priming arises in our 
setup, it should be expected in virtually every microtask and 
crowdsourced workflow where workers are called upon to provide qualitative
judgment.

@@

\section{Prior work}
\subsection{Effects of microtask design on response quality}
% Accuracy, reproducibility, and a controlled task environment are 
% important
Microtask platforms are increasingly used clean and codify datasets,
administer experiments with human participants, and distribute 
surveys.  Accuracy and replicability are crucial in all of these 
applications.

@@

However this approach can \textit{drastically} overestimate $\theta$
\cite{val-thesis}.  As an illustrative example,
suppose that we have sampled two sets of 1000 words from \textit{possibly} 
different distributions, and that we wish to estimate the divergence 
between these distributions.  
It turns out that, if both sets of words were actually drawn from 
an identical Zipf distribution\footnote{The 
  Zipf distribution is a common model for word frequencies 
  \cite{powers1998applications,zipf1949human}:
  \begin{align}
	p(x) = \frac{x^{-1}}{\sum_{n=1}^{\infty}x^{-1}},
	\label{eq:zipf}
  \end{align}
  where $p(x)$ is the probability of the $x$th most common word.
}, the na\"ive approach would typically lead one to report 
$\hat{\theta} \approx 65\%$, event though, in reality $\theta = 0\%$
(this can be shown by simulated sampling).

@@

We will now establish Inequality~\ref{eq:sup:l1} by considering an
optimal classifier having accuracy $\eta_*$.  
Let us assume that workers are shown 
$\mathrm{design}_1$ or $\mathrm{design}_2$ with equal probability.
If the worker gives the response $x$, it is optimal to guess
that the worker was shown the design most likely to elicit $x$.
In other words, if $p_i(x)$ is the probability that a worker shown 
$\mathrm{design}_i$ responds with $x$, then it is optimal to 
guess that the worker saw $\mathrm{design}_j$, where 
$j = \arg\max_j{p_j(x)}$.

@@

\begin{figure}
	\includegraphics[scale=0.75]{figs/theta.pdf}
	\caption{
		Empirical bias, $\theta_\mathrm{NB}$, measured using a na\"ive Bayes 
		classifier, induced in image-labeling tasks, due to workers' 
		exposure to framing and prior tasks.  
		(\textbf{A}) Exposing workers to initial tasks induced more severe
		bias than framing, except when echoed framing was used. 
		(\textbf{B}) As workers proceeded through test tasks, 
		the bias due to initial tasks waned, 
		but remained significant even after five tasks.  
		Standard error bars are shown.
	}
	\label{fig:theta}
\end{figure}

\begin{table}
\small
\centering
\setlength{\tabcolsep}{1pt}
\begin{tabular}{c c c c c}
\toprule
Experiment & \parbox[c]{3.0cm}{\centering{Priming modality}} & Initial Tasks & Frame \\
\midrule
\multirow{2}{*}{\textit{intertask-food-objects}} 
& \multirow{2}{*}{initial tasks} & \textit{food} & none \\
  & & \textit{objects} & none  \\

@@

In our first experiment called \textit{intertask-food-objects},
workers were assigned to either a \textit{food} or \textit{objects} 
treatment.  Workers from these treatments performed initial tasks 
in which they labeled, respectively, 
images depicting either food or (non-food) objects
(see example images in Figure~\ref{fig:task}A and B).  
Workers from both treatments then performed five test tasks, which
contained images
depicting both food and objects (Figure~\ref{fig:task}C).

@@

Having affirmed the \textit{existence} of intertask effects, we sought to
measure the severity of the bias induced, i.e. $\theta$.
To measure $\theta$ we used a na\"ive Bayes classifier based on a 
multinomial distribution over the word frequencies in worker's labels, 
and measured it's accuracy using leave-one-out cross-validation.  
We chose the na\"ive Bayes
classifier for three reasons.  First, it performs well even when the 
number of features is large compared to the number of training examples
\cite{bickel2004,hastie2009elements}.  
Second, there are no hyperparameters to 
optimize, which eliminates the need to partition the response data into
dev and test sets.
Third, the conditional independence assumption, normally 
undertaken for pragmatic reasons \cite{Zhang2004562}, 
is probably a relatively mild in 
relation to image labels, since they are likely to be less dependent
on one another than in a coherent passage of text.  We repeated our
analysis used a support vector machine and found very similar results.

@@

Based on the classifier accuracy, intertask effects in the 
experiment \textit{intertask-food-objects}
lead to (a lower bound) bias of $\theta=30.6\%$ between workers from the 
\textit{food} and \textit{objects} treatments (Figure~\ref{fig:theta}A).
It is difficult to visualize the distributions of word frequencies,
but for reference, the distributions shown in 
Figure~\ref{fig:l1_example}B have $\theta = 30\%$.  This represents a 
substantial potential to distort microtask data.

@@

To assess the persistence of intertask effects, we replicated the 
\textit{intertask-food-objects} experiments, rotating the ordering of the
test tasks, so that each test task occupied each of the possible 5 
test-task positions (see Figure \ref{fig:task-schematic}).  
By taking the average bias induced for each test task when 
occupying a given position, we determined $\theta$ as a function of the 
task position.
Bias was strongest for the first test task (28.1\%), 
but remained significant through all five test tasks 
($\alpha=0.05$) sustaining over 12\% bias even when the initial and test 
tasks were separated by four intervening tasks (Figure~\ref{fig:theta}B).

We conducted variants of these experiments using different images, to see 
whether this trend was robust.  In the experiment 
\textit{intertask-food-culture},
workers were either assigned to a \textit{food} or \textit{culture} treatment.
The initial tasks contained either images 
depicting food (Figure~\ref{fig:task}D), or depictions of cultural scenes 
(of dance, sport, or music) (Figure~\ref{fig:task}E).  The test tasks, 
which 
were, again, identical for both treatments, depicted meals of diverse 
cultural origin
%\footnote{
%  	The food depicted in the initial tasks 
%	in \textit{intertask-food-culture} are arguably still of
%	diverse cultural origin, but this would only tend to make our 
%	results more conservative.
%}
(Figure~\ref{fig:task}F).

@@

To better understand the nature of 
intertask effects, we investigated the vocabulary
that workers used to label test tasks. One might expect that, within a given 
experiment, those workers exposed to food (whether through framing or initial
tasks) would label test tasks using 
food-related words more often.  To test this we developed an 
automated approach to labeling food-words based on the 
WordNet knowledge base \cite{felbaum1998wordnet}.

WordNet provides
hyponym and hypernym relations between words.  A hypernym is a 
generalization (for example, ``bread'' is a hypernym of ``pumpernickel''), 
while a hyponym is a specialization.  We took took the set of food-words
to be all those reachable through
a chain of hyponym relations from the word-senses 
\texttt{food.n.01} and \texttt{food.n.02}, in other words, all words
denoting a specific kind of food (a total of 3590 words).  
To improve the coverage of words for
less widespread foods, we augmented this set with all words
discovered while crawling the World Food section of \url{allrecipes.com} 
which were not already in WordNet.

>>>>

\numberofauthors{3}
\author{%
  \alignauthor{
	Edward Newell \\
    \affaddr{School of Computer Science, McGill University}\\
    \affaddr{Montreal, Qu\'ebec,
	  Canada
	}\\
	\email{edward.newell@mail.mcgill.ca}}\\
  \alignauthor{
	Derek Ruths \\
    \affaddr{School of Computer Science, McGill University}\\
    \affaddr{Montreal, Qu\'ebec,
	  Canada
	}\\
	\email{derek.ruths@mcgill.ca}
  }\\
}

@@

\begin{abstract}
Microtask platforms are becoming commonplace tools for performing human
research, producing gold-standard data, and annotating large datasets.
These platforms connect \textit{requesters}
(researchers or companies) with large populations (crowds) of workers, who 
perform small tasks, typically taking less than five minutes each.
A topic of ongoing research concerns the design of tasks that elicit high
quality annotations.
Here we identify a seemingly banal feature of nearly all crowdsourcing 
workflows that profoundly impacts workers' responses.
Microtask assignments typically consist of a sequence 
of tasks sharing a common format (e.g., circle galaxies in an image). 
Using image-labeling, a canonical microtask format, we 
discover that earlier tasks have a priming effect on the worker, shifting 
the distribution of future responses by 30-50\% 
(total variational distance). 
Specifically, prior tasks influence the content that workers focus on, 
as well as the richness and specialization of responses. 
We call this phenomenon \textit{intertask effects}.
We compare intertask effects to the overt framing, 
effected by stating the requester's research interest, 
and find that intertask effects are on par or stronger.
While intertask effects can be a source of systematic bias, 
our results suggest that, with appropriate task design, 
they might be leveraged to hone worker focus and acuity, 
helping to elicit reproducible, expert-level judgments.
Intertask effects are a crucial aspect of human computation that should be
considered in the design of any crowdsourced study.
\end{abstract}

@@

\section{Introduction}
There are many tasks that are trivial for people, but difficult to solve
programmatically.  
Such tasks include tagging and categorizing images,
%\cite{6116320,Zhai2012357}, 
coding and transcribing media,
%\cite{chandler2013breaking,Berinsky2012351,Finnerty2013},%paolacci2010running},
%judging the relevancy or quality of content,
and performing surveys for 
academic or market research purposes (see Table~\ref{table:task_composition} for a listing of task types seen on the Amazon's Mechanical Turk 
microtask platform).
%\cite{le2010ensuring,grady2010crowdsourcing,alonso2009can,kazai2013analysis}.
Many tasks are ill-defined, in the sense that they do not have
a clear ``correct'' response, and require high-level, qualitative judgment.
Microtask platforms are marketplaces that help fill the gap in current 
computational capabilities, by matching requesters, who need to have such 
tasks completed, with human workers.  Amazon's Mechanical Turk\footnote{\url{mturk.com}}, and 
CrowdFlower\footnote{\url{crowdflower.com}} are two popular microtask 
platforms.

@@

To investigate the phenomenon in a
highly generalizable setting, we sought a task that would involve 
qualitative judgment and enable relatively unconstrained responses. 
We also sought a task that would be a typical exemplar of the kinds of
tasks actually seen on microtask platforms (see 
Table~\ref{table:task_composition} for examples).  For this purpose, we 
adopted
image-labeling as a canonical microtask, in which workers provide 
descriptive labels to images using free-text inputs.  Image labeling is
a qualitative task without clearly ``correct'' or ``incorrect'' responses,
and is among the most common kinds of tasks on the major crowdsourcing
platform Amazon's Mechanical Turk (see Table \ref{table:task_composition}).  If task-induced priming arises in our 
setup, it should be expected in an overwhelming number of microtasks and 
crowdsourced workflows where workers are called upon to provide qualitative
judgment.

@@

\section{Prior work}
\subsection{Effects of microtask design on response quality}
% Accuracy, reproducibility, and a controlled task environment are 
% important
Microtask platforms are increasingly used to clean and codify datasets,
administer experiments with human participants, and distribute 
surveys.  Accuracy and replicability are crucial in all of these 
applications.

@@

However this approach can \textit{drastically} overestimate $\theta$
\cite{val-thesis}.  As an illustrative example,
suppose that we have sampled two sets of 1000 words from \textit{possibly} 
different distributions, and that we wish to estimate the divergence 
between these distributions.  
It turns out that, if both sets of words were actually drawn from 
an identical Zipf distribution\footnote{The 
  Zipf distribution is a common model for word frequencies 
  \cite{powers1998applications,zipf1949human}:
  \begin{align}
	p(x) = \frac{x^{-1}}{\sum_{n=1}^{\infty}x^{-1}},
	\label{eq:zipf}
  \end{align}
  where $p(x)$ is the probability of the $x$th most common word.
}, the na\"ive approach would typically lead one to report 
$\hat{\theta} \approx 65\%$, even though, in reality $\theta = 0\%$
(this can be shown by simulated sampling).

@@

We can establish Inequality~\ref{eq:sup:l1} by considering an
optimal classifier having accuracy $\eta_*$.  
Let us assume that workers are shown 
$\mathrm{design}_1$ or $\mathrm{design}_2$ with equal probability.
If the worker gives the response $x$, it is optimal to guess
that the worker was shown the design most likely to elicit $x$.
In other words, if $p_i(x)$ is the probability that a worker shown 
$\mathrm{design}_i$ responds with $x$, then it is optimal to 
guess that the worker saw $\mathrm{design}_j$, where 
$j = \arg\max_j{p_j(x)}$.

@@

\begin{figure}[t]
	\includegraphics[scale=0.75]{figs/theta.pdf}
	\caption{
		Empirical bias, $\theta_\mathrm{NB}$, measured using a na\"ive Bayes 
		classifier, induced in image-labeling tasks, due to workers' 
		exposure to framing and prior tasks.  
		(\textbf{A}) Exposing workers to initial tasks induced more severe
		bias than framing, except when echoed framing was used. 
		(\textbf{B}) As workers proceeded through test tasks, 
		the bias due to initial tasks waned, 
		but remained significant even after five tasks.  
		Standard error bars are shown.
	}
	\label{fig:theta}
\end{figure}

\begin{table}[b!]
\small
\centering
\setlength{\tabcolsep}{1pt}
\begin{tabular}{c c c c c}
\toprule
Experiment & \parbox[c]{3.0cm}{\centering{Priming modality}} & Initial Tasks & Frame \\
\midrule
\multirow{2}{*}{\textit{intertask-food-objects}} 
& \multirow{2}{*}{initial tasks} & \textit{food} & none \\
  & & \textit{objects} & none  \\

@@

{\bf Existence of inter-task effects.} In our first experiment called \textit{intertask-food-objects},
workers were assigned to either a \textit{food} or \textit{objects} 
treatment.  Workers from these treatments performed initial tasks 
in which they labeled, respectively, 
images depicting either food or (non-food) objects
(see example images in Figure~\ref{fig:task}A and B).  
Workers from both treatments then performed five test tasks, which
contained images
depicting both food and objects (Figure~\ref{fig:task}C).

@@

{\bf Strength of inter-task effects.} Having affirmed the \textit{existence} of intertask effects, we sought to
measure the severity of the bias induced, i.e. $\theta$.
To measure $\theta$ we used a na\"ive Bayes classifier based on a 
multinomial distribution over the word frequencies in worker's labels, 
and measured it's accuracy using leave-one-out cross-validation.  
We chose the na\"ive Bayes
classifier for three reasons.  First, it performs well even when the 
number of features is large compared to the number of training examples
\cite{bickel2004,hastie2009elements}.  
Second, there are no hyperparameters to 
optimize, which eliminates the need to partition the response data into
dev and test sets.
Third, the conditional independence assumption, normally 
undertaken for pragmatic reasons \cite{Zhang2004562}, 
is probably a relatively mild in 
relation to image labels, since they are likely to be less dependent
on one another than in a coherent passage of text.  We repeated our
analysis used a support vector machine and found very similar results.

@@

Based on the classifier accuracy, intertask effects in the 
experiment \textit{intertask-food-objects}
lead to (a lower bound) bias of $\theta=30.6\%$ between workers from the 
\textit{food} and \textit{objects} treatments (Figure~\ref{fig:theta}A).
It is difficult to visualize the distributions of word frequencies,
but for reference, the distributions shown in 
Figure~\ref{fig:l1_example}B have $\theta = 30\%$.  This represents a 
substantial potential to distort microtask results.

@@

{\bf Persistence of inter-task effects.} To assess the persistence of intertask effects, we replicated the 
\textit{intertask-food-objects} experiments, rotating the ordering of the
test tasks, so that each test task occupied each of the possible 5 
test-task positions (see Figure \ref{fig:task-schematic}).  
By taking the average bias induced for each test task when 
occupying a given position, we determined $\theta$ as a function of the 
task position.
Bias was strongest for the first test task (28.1\%), 
but remained significant through all five test tasks 
($\alpha=0.05$) sustaining over 12\% bias even when the initial and test 
tasks were separated by four intervening tasks (Figure~\ref{fig:theta}B).

{\bf Different tasks, similar effects.} We conducted variants of these experiments using different images, to see 
whether this trend was robust.  In the experiment 
\textit{intertask-food-culture},
workers were either assigned to a \textit{food} or \textit{culture} treatment.
The initial tasks contained either images 
depicting food (Figure~\ref{fig:task}D), or depictions of cultural scenes 
(of dance, sport, or music) (Figure~\ref{fig:task}E).  The test tasks, 
which 
were, again, identical for both treatments, depicted meals of diverse 
cultural origin
%\footnote{
%  	The food depicted in the initial tasks 
%	in \textit{intertask-food-culture} are arguably still of
%	diverse cultural origin, but this would only tend to make our 
%	results more conservative.
%}
(Figure~\ref{fig:task}F).

@@

{\bf Inter-task effects and concept activation.} To better understand the nature of 
intertask effects, we investigated the vocabulary
that workers used to label test tasks. One might expect that, within a given 
experiment, those workers exposed to food (whether through framing or initial
tasks) would label test tasks using 
food-related words more often.  To test this we developed an 
automated approach to labeling food-words based on the 
WordNet knowledge base \cite{felbaum1998wordnet}.

WordNet provides
hyponym and hypernym relations between words.  A hypernym is a 
generalization (for example, ``bread'' is a hypernym of ``pumpernickel''), 
while a hyponym is a specialization.  We took the set of food-words
to be all those reachable through
a chain of hyponym relations from the word-senses 
\texttt{food.n.01} and \texttt{food.n.02}, in other words, all words
denoting a specific kind of food (a total of 3590 words).  
To improve the coverage of words for
less widespread foods, we augmented this set with all words
discovered while crawling the World Food section of \url{allrecipes.com} 
which were not already in WordNet.
